---
title: "Survival analysis"
author: "yacho"
date: "2023-01-10"
output: html_document
---
```{r}
# firstly load required package, if not those pacakges not installed before, then
#install.packages('survival')# can change the pacakge name inside the bracket
#install.packages('survival')
#install.packages("ggfortify")
#install.packages("ggsurvfit")
library(survival)
library(survminer)
library(ggplot2)
library(ggfortify)
library(plotly)
library(ggsurvfit)
```

# The first example is analysis the whether there is significant difference in two different drugs with conditions as cancer, the reading data is already extracted as the standard format used for survival analysis used for check the correctness of Kaplan Meier curve plot by Tableau


```{r}
data = read.csv(file = 'preprocessed_data_Tableau/two_drug.csv')
names(data)[names(data) == 'medication'] <- 'group'
data[order(data$group,data$Time),]
# for kaplan Meier analysis, the standard format need with column
#Time: time to events
#Status about whether the events happened or not 0:not happened or censored data 1:events happened
#medication: group about the medication the patient used 
```


# next part there are two function can used for plot Kaplan Meier curve with standard format
```{r}
# this function need input with dataframe and a string state the duration unit
get_kaplan_meier_plot1 <- function(data,time_unit){
  data = data.frame(Time = data$Time, Status = data$Status,group = data$group)
  km_fit <- surv_fit(Surv(Time, Status) ~ group, data=data)
  # if there is exactly two group then we can use the logrank test 
  whetherlogranktest = (length(unique(data$group))==2)
  pvalue_text = ""
  if(whetherlogranktest){
    pvalue = toString(round(survdiff(Surv(Time, Status) ~ group, data=data)$pvalue,8))
    pvalue_text = paste0("logrank pvalue: \n ",pvalue)
  }
  
  p = autoplot(km_fit) + 
   labs(x = paste("\n Survival Time in", time_unit) , y = "Survival Probabilities \n", 
   title = "Kaplan Meier plot") + 
   ylim(0, 1) +
    annotate("text", x=max(data$Time)/5, y=0, label= pvalue_text)+

   theme(plot.title = element_text(hjust = 0.5), 
   axis.title.x = element_text(face="bold", colour="#FF7A33", size = 12),
   axis.title.y = element_text(face="bold", colour="#FF7A33", size = 12),
   legend.title = element_text(face="bold", size = 10))

  
  # ggplot do not developed for CI
  ggplotly(p)
}

get_kaplan_meier_plot2 <- function(data,time_unit){
  data = data.frame(Time = data$Time, Status = data$Status,group = data$group)
  km_fit <- surv_fit(Surv(Time, Status) ~ group, data=data)
  p = ggsurvplot(km_fit, 
           conf.int = TRUE, 
           risk.table = "abs_pct", 
           pval = TRUE,
           pval.method = TRUE,
           risk.table.height = 0.25, # avoid overlap the table
  ggtheme = theme_bw()  )
  (p)
  ggplotly(p[[1]])
  
}




```


```{r}
length(unique(data$group))
data
time_unit = 'Day'
get_kaplan_meier_plot2(data,time_unit)
get_kaplan_meier_plot1(data,time_unit)
```







```{r}

data =  read.csv(file = 'preprocessed_data_Tableau/all_die_1_year.csv')
names(data)[names(data) == 'DESCRIPTION'] <- 'group'
get_kaplan_meier_plot1(data,time_unit)

```


```{r}
data =  read.csv(file = 'preprocessed_data_Tableau/all_alive.csv')
names(data)[names(data) == 'DESCRIPTION'] <- 'group'
get_kaplan_meier_plot1(data,time_unit)

```

```{r}
data =  read.csv(file = 'preprocessed_data_Tableau/all_die_1_day.csv')
names(data)[names(data) == 'DESCRIPTION'] <- 'group'
get_kaplan_meier_plot1(data,time_unit)
```

```{r}
data =  read.csv(file = 'preprocessed_data_Tableau/all_die_prop_time.csv')
names(data)[names(data) == 'DESCRIPTION'] <- 'group'
get_kaplan_meier_plot1(data,time_unit)
```

```{r}
data =  read.csv(file = 'preprocessed_data_Tableau/two_condition.csv')
names(data)[names(data) == 'DESCRIPTION'] <- 'group'
get_kaplan_meier_plot1(data,time_unit)
```



# After getting the plot, we also interested in whether the two group statistical significant or not, hence we need to do the logrank test

```{r}

data = read.csv(file = 'preprocessed_data_Tableau/two_drug.csv')
names(data)[names(data) == 'medication'] <- 'group'
survdiff(Surv(Time, Status) ~ group, data=data)$pvalue
```




# Next is about the cox model 

# if there is one variable and is binary classes, we can find out the pvalue of the logranktest value is same as the Score(logrank) test in the summary of cox model
```{r}
data = read.csv(file = 'preprocessed_data_Tableau/two_drug.csv')
coxfit = coxph(Surv(Time, Status) ~ medication, data=data)
summary(coxfit)$sctest[3]
survdiff(Surv(Time, Status) ~ medication , data=cox_info)$pvalue
```


# for multiple variable and with continuous variables

```{r}
data = read.csv(file = 'preprocessed_data_Tableau/two_drug.csv')
coxfit = coxph(Surv(Time, Status) ~ medication + age.cancer., data=data)
summary(coxfit)

```

# To interpret this summary of the coxfit: 
there is a 0.78668 unit decrease in expected log of the relative hazard for Paclitaxel 100 MG Injection as compared to 100 ML Epirubicin Hydrochloride 2 MG/ML Injection, holding other factor constant.

there is a 0.02420 unit decrease in expected log of the relative hazard for unit increased in age, holding other factor constant.


